---
title: "Reproducible Research Project 1"
---
## Activity Data Writeup

In this project we have data from 1 activity monitoring device from one anonymous person. We will be looking at their steps taken per day and month over the months of October and November 2012.

First we will read in our data set and remove NA values. In addition, I am 
going to reformat the date column into POSIX format.

```{r setup, results = 'hide'}
data <- read.csv("activity.csv")
cleandata <- data[!is.na(data$steps),]
cleandata$date <- as.POSIXct(cleandata$date, format="%Y-%m-%d")
```

A histogram of the total number of steps taken each day. 
```{r histogram}
perday <- aggregate(cleandata$steps, by=list(cleandata$date), FUN=sum)
colnames(perday) <- c("date","steps")
hist(c(perday$steps,perday$date), main = "Count of Steps per Day", 
     xlab = "Steps Taken", ylab = "Number of Days Recorded",
     breaks = 5)
```

Next I have determined the mean and median of the number of steps taken per day.
```{r mean and median}
mean(perday$steps, na.rm = TRUE)
median(perday$steps, na.rm = TRUE)
```

Here I have made a time series plot of the average number of steps taken over
the 5 minute intervals that data was collected throughout the day.
```{r timeseries}
intervals <- tapply(cleandata$steps, cleandata$interval, mean)
plot(as.numeric(names(intervals)), intervals, type = "l")
```

Now I will find the 5 minute interval that our particpant took the most steps
throughout the day. 
```{r max interval}
avginterval <- aggregate(cleandata$steps ~ cleandata$interval, FUN = sum)
maxsteps <- sort(avginterval$`cleandata$steps`, decreasing = TRUE)
print(avginterval[avginterval$`cleandata$steps` == maxsteps[1],])
```

